On the presentation of orthogonal groups over polynomial rings

### 2019-01-30 dl-paper.tex
Thorough rewrite

\title{On the presentation of orthogonal groups over polynomial rings}
\keywords {Steinberg group, $K_2$-functor, Serre problem
 {\em Mathematical Subject Classification (2010):} 19C20}
\author {Sergey Sinchuk}
\address{Chebyshev laboratory, St. Petersburg State University, St. Petersburg, Russia}
\email {sinchukss at gmail.com}
\date {\today}

@@

\subsection{Steinberg groups}
Let $\Phi$ be a reduced and irreducible root system of rank $\geq 2$ and $R$ be a commutative ring with $1$. Recall that in this case the \emph{Steinberg group} $\St(\Phi, R)$ can be defined by means of generators $x_{\alpha}(\xi)$ where $\xi\in R, \alpha\in\Phi$ and relations:
\begin{align}
& \phantom{[}
x_\alpha(s) x_\alpha(t) = x_\alpha(s+t),\ \alpha\in\Phi,\ s,t\in R; \label{rel:add}\\
& [x_\alpha(s), x_\beta(t)] = \prod\limits_{i,j\in\mathbb{N}}
 x_{i\alpha + j\beta}\left(N_{\alpha\beta ij}\, s^i t^j\right),\quad \alpha,\beta\in\Phi,\ \alpha\neq\pm\beta,\ s,t\in R. \label{rel:CCF}
\end{align}
The indices $i$, $j$ appearing in the right-hand side of the above relation range over
all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The structure constants $N_{\alpha \beta i j}=\pm 1,2,3$ appearing in \eqref{rel:CCF} depend only on $\Phi$ and can be computed explicitly.

@@

\begin{comment}
Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$ 
\end{comment}
\subsection{Unstable Quillen K-groups}
\begin{comment}
In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. 
Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).
\end{comment}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following reqirements:
\begin{enumerate}
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

%TODO: Give the definition of $D_{R, I}$ and of 
\begin{prop}\label{characterization}
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is a universal relative central extension of 
    $S(R) \to S(R/I)$, whose kernel is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is an exact sequence which is natural in $(R, I)$:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \Ker(\mu).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook} (here we use~\eqref{req:subfunc}--\eqref{req:uce}). %provided one verifies that all the intermediate isomorphisms are natural in $R$.

@@

In the above definition we could have replaced the orthogonal group with a different isogeneous form of the orthogonal group, e.\,g. 
 we could have chosen $G_n = \mathrm{SO}_{2n}$ or $G_n = \mathrm{Spin}_{2n}$ and then defined $P_n$ as its elementary subfunctor.
Of course, this would change the functors $\K_i(G_n, P_n, -)$ for $i=1,2$ however, as the proof of~\cref{characterization} suggests
 it would not change the functors $\K_i(G_n, P_n, -)$ for $i\geq 3$.

Now let $G_n, P_n$ be again as in~\cref{ex:orth}. 
Let us formulate a stability result for orthogonal Quillen K-groups (see~\cite[Theorem~9.4]{Pa89}).
\begin{theorem}[Panin] \label{lem:Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain. Set $a = 1,2$ or $3$ in each of these respective cases.
 Then the canonical map $\K_i(G_n, P_n, R) \to \K_i(G_{n+1}, P_{n+1}, R)$ is an epimorphism for $n \geq b$ 
 and an isomorphism for $\ell \geq b + 1$, where $b = \mathrm{max}(2i, a+i-1)$. \end{theorem}

Of course, in the limit, the groups $\K_i(\mathrm{O}_\infty, \mathrm{EO}_\infty, R) = KO_i(R)$ coincide with Karoubi's hermitian groups ${}_{-1}\!L_i(R)$ defined in~\cite{Ka80}.
The modern version of hermitian K-theory is the theory of higher Grothendieck--Witt groups of schemes.
We refer the reader to~\cite[\S~2]{AF17} or~\cite[\S~2]{FRS12} for a more detailed introduction to it.
For our purposes it would be sufficient to think of $\GW_i^{[k]}(R)$ as a shorthand for the following:
\begin{equation}
 \GW_i^{[k]}(R) = \left\{\begin{array}{ll} KO_i(R), & k = 0, \\ U_i(R), & k = 1, \\ KSp_i(R), & k = 2, \\ {}_{-1}\!U_i(R), & k = 3. \end{array}\right.
\end{equation}
The notation for the upper index $[k]$ suggests that Grothendieck--Witt groups are $4$-periodic in $k$.
The groups ${}_{\pm 1}\!U_i(R)$ are defined in~\cite{Ka80} and we will not use their definition.

In the sequel we will need the following result.
\begin{theorem}[Bass Fundamental Theorem]\label{bass-ft} Suppose that $2 \in R^*$, then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ there is a split exact sequence of abelian groups
 \[ \xymatrix{ 0 \ar[r] & \GW_i^{[k]}(R) \ar[r] &  \GW_i^{[k]}(R[X, X^{-1}]) \ar[r]  & \GW_{i-1}^{[k-1]}(R) \ar[r] & 0.} \] \end{theorem}
\begin{proof} This is a special case of~\cite[Theorem~9.13]{Sch16} and the remark that follows it. \end{proof}

\subsection{Relative Steinberg groups}

The aim of this section is to briefly remind the reader definitions of relative Steinberg groups and their basic properties. We refer the reader to~\cite[Section~3]{S15} for a more detailed exposition.
%TODO: Define the category of pairs (so that all constructions below become functorial wrt it).
...

%Thus, by our definition, the relative Steinberg group is a central extension of $\St(\Phi, R)$.
\begin{equation}\label{eq:relative-St}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r] & \St(\Phi, R, I) \ar[r] & \St(\Phi, R) \ar[r]^{\pi^*} & \St(\Phi, R/I) \ar[r] & 1 }
\end{equation}
It is not hard to show that $C(\Phi, R, I) = 0$ in the special case when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits, see~\cite[Lemma~8]{S15}.

Unstable relative Steinberg groups share many good properties with their stable counterparts studied in~\cite{Lo78} provided the following assumptions on $\Phi$ hold: \begin{enumerate}
 \item $\St(\Phi, R)$ is perfect and super-perfect for all commutative $R$ (i.\,e. the homology groups $\HH_i(\St(\Phi, R), \ZZ)$ vanish for $i=1,2$).
 \item for all commutative rings $R$ the group $\K_2(\Phi, R)$ is central in $\St(\Phi, R)$.
\end{enumerate}
For example, under these assumptions it can be shown that the map $\St(\Phi, R, I) \to \St(\Phi, R)$ is a universal relative central extension of a pair $(\St(\Phi, R/I), \St(\Phi, R))$ in the sense of~\cite{Lo78} (the proof is identical to~\cite[Corollary~3]{S15}).

The following result was proved in \cite[Proposition~3.2]{Ste13}.
\begin{lemma}\label{lem:Zgen} Let $\Phi$ be an irreducible root system of rank $\geq 2$. 
Let $G$ denote the normal closure in $\St(\Phi, R)$ of the subgroup spanned by elementary root unipotents of level $I$, i.\,e. $G=\langle x_\alpha(s) \mid \alpha\in\Phi,\ s\in I\rangle^{\St(\Phi, R)}$.

@@

\section{Proof of the main result}
\subsection{Overview of the proof}
\subsection{First injectivity theorem}
For a commutative local ring $A$ with the maximal ideal $M$ and residue field $k$
we denote by $B = B_{A, M}$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$.
It is clear that $I := M[X, X^{-1}]$ can be considered as an ideal of both $B$ and $R$.
\begin{lemma}\label{lem:lemma32} Let $\Phi$ be a simply-laced root system of rank $\geq 3$.
Consider the following commutative square of canonical maps.
\[ \xymatrix{
    \St(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B) \ar[d] \\
    \St(\Phi, R, I) \ar[r] \ar@{-->}[ur] & \St(\Phi, R)
   } \]
Then there exists a diagonal arrow which makes the diagram commute.   
\end{lemma} 
\begin{proof}
 Notice that $R$ is isomorphic to the principal localisation of $B$ at $X$
  and that $I$ is uniquely $X$-divisible in the sense of~\cite[\S~4]{LS17}.
 Thus, in the special case $\Phi = \rA_3$ the assertion of the lemma follows from~\cite[Theorem~3]{LS17}.

 Now the general case can be obtained as a corollary of amalgamation theorem~\cite[Theorem~9]{S15}
  (cf.~\cite[\S~4]{LS17} for more details).
 %TODO: Also give reference on a similar result of Lavrenov
 %TODO: Rethink the proof of amalgamation theorem...
\end{proof}

@@

\begin{corollary} \label{field-injectivity} For $\Phi\neq\rG_2$ the map $\St(\Phi, k[X]) \to \St(\Phi, k[X, X^{-1}])$ is injective. \end{corollary}
\begin{proof} The result follows from consideration of the commutative diagram
\[\xymatrix{ & \K_2(\Phi, k) \ar[dl]_{\cong} \ar@{^{(}->}[dr] & \\
               \K_2(\Phi, k[X]) \ar[rr] &               & \K_2(\Phi, k[X, X^{-1}])} \]
in which the left arrow is an isomorphism by the corollary of~\cite[Satz~1]{Re75} and the right arrow is split injective by the above lemma. \end{proof}

\begin{prop}
 Suppose that the canonical map $C(\Phi, B, I) \to C(\Phi, R, I)$ is surjective,
  and that Tulenbaev lifting property holds for $\St(\Phi, R)$. 
 Then the canonical map of Steinberg groups $\St(\Phi, B) \to \St(\Phi, R)$ is injective.
\end{prop}
\begin{proof}
 Consider the following commutative diagram with exact rows: 
\begin{equation*} \xymatrix{
 C(\Phi, B, I) \ar[r]^{j_B} \ar@{->>}[d]_{f} & \St(\Phi, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\Phi, B) \ar[r]^{\pi_B} \ar[d]_{h} & \St(\Phi, k[X]) \ar[d]_{i} \\
 C(\Phi, R, I) \ar[r]^{j_R}         & \St(\Phi, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\Phi, R) \ar[r]^{\pi_R}        & \St(\Phi, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$ also lies in $\Ker(\pi_B)$ and hence comes from some $b \in \St(\Phi, B, I)$ via $\mu_B$. Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\Phi, R, I)$ such that $j_R(c) = g(b)$. By our assumption, $f$ is surjective, therefore
 $c = f(d)$ for some $d \in C(\Phi, R, I)$ and we obtain the required assertion:
 \[ 1 = \mu_Bj_B(d) = tgj_B(d) = tj_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]

@@

\begin{theorem} Suppose $A$ is a local ring of characteristic $\neq 2$, then for $\ell \geq 7$
  the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}
 In view of the above proposition and~\cref{lem:lemma32} we are left to verify that the map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective. Consider the following commutative diagram, whose rows are given by \cref{characterization}:
 \begin{equation*}\xymatrix{
  \K_3(\rD_\ell, B) \ar[r] \ar[d] & \K_3(\rD_\ell, k[X]) \ar[d] \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
  \K_3(\rD_\ell, R) \ar[r]        & \K_3(\rD_\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
 Invoking~\cref{bass-ft} and \cite[Lemma~2.2]{FRS12} we obtain that $\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k)$.
 In view of~\cref{lem:Panin-stability} the central vertical map is isomorphic to the canonical map
 $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$ and therefore is surjective. The requred surjectivity of $f$ now follows.
\end{proof}

>>>>

\title{On the presentation of orthogonal groups over polynomial rings}
\keywords {Steinberg group, $K_2$-functor, Serre problem
 {\em Mathematical Subject Classification (2010):} 19C20}
\author{Andrei Lavrenov}
\address{Mathematisches Institut der Universit\"at M\"unchen, Theresienstr. 39, D-80333 M\"unchen}
\email{avlavrenov at gmail.com}

\author {Sergey Sinchuk}
\address{Chebyshev laboratory, St. Petersburg State University, St. Petersburg, Russia}
\email {sinchukss at gmail.com}
\date {\today}

@@

\subsection{Steinberg groups}
%TODO: Describe what root systems are sympelctic and which are not
Let $\Phi$ be a reduced and irreducible root system of rank $\geq 2$ and $R$ be a commutative ring with $1$. Recall that in this case the \emph{Steinberg group} $\St(\Phi, R)$ can be defined by means of generators $x_{\alpha}(\xi)$ where $\xi\in R, \alpha\in\Phi$ and relations:
\begin{align}
& \phantom{[}
x_\alpha(s) x_\alpha(t) = x_\alpha(s+t),\ \alpha\in\Phi,\ s,t\in R; \label{rel:add}\\
& [x_\alpha(s), x_\beta(t)] = \prod\limits_{i,j\in\mathbb{N}}
 x_{i\alpha + j\beta}\left(N_{\alpha\beta ij}\, s^i t^j\right),\quad \alpha,\beta\in\Phi,\ \alpha\neq\pm\beta,\ s,t\in R. \label{rel:CCF}
\end{align}
The indices $i$, $j$ appearing in the right-hand side of the above relation range over
all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The structure constants $N_{\alpha \beta i j}=\pm 1,2,3$ appearing in \eqref{rel:CCF} depend only on $\Phi$ and can be computed explicitly.

@@

\begin{comment}
Suppose for a moment that $\langle \alpha, \beta \rangle = -1$ and  $\langle \beta, \alpha \rangle = -1$ then
\[ \{s, t^{-1} \} = \{s,  t^{-1}\}_\alpha = \{t, s^{-1} \}_\beta^{-1} = \{s^{-1}, t\} \]
In particular, $\{s, s^{-1}\} = \{s, s^{-1}\}^{-1}$ 
\end{comment}
\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following reqirements:
\begin{enumerate}
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

%TODO: Give the definition of $D_{R, I}$ and of the category of Pairs
\begin{prop}\label{characterization}
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is the universal relative central extension of 
    the map $S(R) \to S(R/I)$. The kernel of $\mu$ is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is a natural exact sequence:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \Ker(\mu).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook} (here we use~\eqref{req:subfunc}--\eqref{req:uce}). %provided one verifies that all the intermediate isomorphisms are natural in $R$.

@@

In the above definition we could have replaced the group $\mathrm{O}_{2n}$ with a different isogeneous form of the orthogonal group, e.\,g. 
 we could have chosen $G_n = \mathrm{SO}_{2n}$ or $G_n = \mathrm{Spin}_{2n}$ and then have defined $P_n$ as its elementary subfunctor.
Of course, this would change the functors $\K_i(G_n, P_n, -)$ for $i=1,2$ however, as the proof of~\cref{characterization} suggests %TODO: Does it really change K_2?
 it would not change $\K_i(G_n, P_n, -)$ for $i\geq 3$.

Now let $G_n, P_n$ be again as in~\cref{ex:orth}. We write $KO_i(n, R)$ as a shorthand for $\K_i(G_n, P_n, R)$.
Let us formulate the stability theorem for orthogonal Quillen K-groups (see~\cite[Theorem~9.4]{Pa89}).
\begin{theorem}[Panin] \label{lem:Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain. Set $a = 1,2$ or $3$ in each of these respective cases.
 Then the canonical map $KO_i(2n, R) \to KO_i(2(n+1), R)$ is an epimorphism for $n \geq b$ 
 and an isomorphism for $\ell \geq b + 1$, where $b = \mathrm{max}(2i, a+i-1)$. \end{theorem}

Of course, the limit groups $\K_i(\mathrm{O}_\infty, \mathrm{EO}_\infty, R) = KO_i(R)$ are exactly hermitian K-groups ${}_{-1}\!L_i(R)$ defined in~\cite{Ka80}.
There is a modern version of hermitian K-theory, namely the theory of higher Grothendieck--Witt groups of schemes.
We refer the reader to~\cite[\S~2]{AF17} or~\cite[\S~2]{FRS12} for an introduction to this theory.
For our purposes it would be sufficient to restrict our attention to to the affine case and to think of $\GW_i^{[k]}(R)$ as a shorthand for
\begin{equation}
 \GW_i^{[k]}(R) = \left\{\begin{array}{ll} KO_i(R), & k = 0, \\ U_i(R), & k = 1, \\ KSp_i(R), & k = 2, \\ {}_{-1}\!U_i(R), & k = 3. \end{array}\right.
\end{equation} %TODO: Reference?
The groups ${}_{\pm 1}\!U_i(R)$ are defined in~\cite{Ka80} and we will not use their definition.
The notation for the upper index $[k]$ suggests that Grothendieck--Witt groups are $4$-periodic in $k$.

In the sequel we will need the following result.
\begin{theorem}[Bass Fundamental Theorem]\label{bass-ft} Suppose that $2 \in R^*$, then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ there is a natural split exact sequence of abelian groups
 \[ \xymatrix{ 0 \ar[r] & \GW_i^{[k]}(R) \ar[r] & \GW_i^{[k]}(R[X, X^{-1}]) \ar[r]  & \GW_{i-1}^{[k-1]}(R) \ar[r] & 0.} \] \end{theorem}
\begin{proof} This is a special case of~\cite[Theorem~9.13]{Sch16} and the remark that follows it. \end{proof}

\subsection{Relative Steinberg groups}
Let us briefly recall the definition and basic properties of relative Steinberg groups.
We refer the reader to~\cite[Section~3]{S15} for a more detailed exposition.
By definition, the relative Steinberg group functor $\St(\Phi, R, I)$ is obtained from the functor $\St(\Phi, R)$
 via the relativization procedure described in~\cref{sec:quillen}.

\begin{lemma}\label{lem:lemma32} Let $\Phi$ be a simply-laced root system of rank $\geq 3$,
$A$ be arbitrary commutative local ring with maximal ideal $M$ and residue field $k$.
Denote by $B = B_{A, M}$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$ and
by $I$ the ideal $M[X, X^{-1}]$ of $B$ (it is clear that $I$ is also an ideal of $R$).
Consider the following commutative square of canonical maps.
\[ \xymatrix{
    \St(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B) \ar[d] \\
    \St(\Phi, R, I) \ar[r] \ar@{-->}[ur] & \St(\Phi, R)
   } \]
Then there exists a diagonal arrow which makes the diagram commute.   
\end{lemma} 
\begin{proof}
 Notice that $R$ is isomorphic to the principal localisation of $B$ at $X$
  and that $I$ is uniquely $X$-divisible in the sense of~\cite[\S~4]{LS17}.
 Thus, in the special case $\Phi = \rA_3$ the assertion of the lemma follows from~\cite[Theorem~3]{LS17}.

 Now the general case can be obtained as a corollary of amalgamation theorem~\cite[Theorem~9]{S15}
  (cf.~\cite[\S~4]{LS17} for more details).
\end{proof}
%TODO: Give reference on a similar result of Lavrenov?

If we denote the kernel of the natural transformation $\mu$ from the definition of $\St(\Phi, R, I)$ by $C(\Phi, R, I)$ we obtain a natural exact sequence:
\begin{equation*}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r]^{j} & \St(\Phi, R, I) \ar[r]^{\mu} & \St(\Phi, R) \ar[r]^{\pi} & \St(\Phi, R/I) \ar[r] & 1.}
\end{equation*}
The group $C(\Phi, R, I)$ is trivial when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits.
This immediately follows e.\,g. from the exact sequence of~\cref{characterization}, it is also possible to give a direct proof of this fact, see \cite[Lemma~8]{S15}.
The following result plays a key role in the proof of~\cref{thm41}.
\begin{lemma} \label{lem:prop41}
In the notation of~\cref{lem:lemma32} assume additionally that the residue field $k$ is of characteristic $\neq 2$.
Then the canonical map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective for $\ell \geq 7$.
\end{lemma}
\begin{proof}
  Consider the following commutative diagram with rows obtained from~\cref{characterization}:
\begin{equation*}\xymatrix{
 KO_3(2\ell, B) \ar[r] \ar[d] & KO_3(2\ell, k[X]) \ar[d]_{f'} \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
 KO_3(2\ell, R) \ar[r]        & KO_3(2\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
In view of~\cref{lem:Panin-stability} the central vertical map $f'$ is isomorphic to the canonical map $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$.
Invoking~\cref{bass-ft} we obtain that \[\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k) \oplus \GW_2^{[3]}(k).\]
By \cite[Lemma~2.2]{FRS12} the group $\GW_2^{[3]}(k)$ is trivial, therefore $f'$ (and hence $f$) are surjective.
\end{proof}

\subsection{Miscellanea}
The following result was proved in \cite[Proposition~3.2]{Ste13}.
\begin{lemma}\label{lem:Zgen} Let $\Phi$ be an irreducible root system of rank $\geq 2$. 
Let $G$ denote the normal closure in $\St(\Phi, R)$ of the subgroup spanned by elementary root unipotents of level $I$, i.\,e. $G=\langle x_\alpha(s) \mid \alpha\in\Phi,\ s\in I\rangle^{\St(\Phi, R)}$.

@@

\section{Proof of the main result}
\subsection{Overview of the proof}
\subsection{First injectivity theorem}

@@

\begin{corollary} \label{field-injectivity} For $\Phi\neq\rG_2$ the map $\St(\Phi, k[X]) \to \St(\Phi, k[X, X^{-1}])$ is injective. \end{corollary}
\begin{proof} The result follows from consideration of the commutative diagram
\[\xymatrix{ & \K_2(\Phi, k) \ar[dl]_{\cong} \ar@{^{(}->}[dr] & \\
               \K_2(\Phi, k[X]) \ar[rr] &               & \K_2(\Phi, k[X, X^{-1}])} \]
in which the left arrow is an isomorphism by the corollary of~\cite[Satz~1]{Re75} and the right arrow is split injective by the above lemma. \end{proof}
\begin{theorem} \label{thm41}
Assume the notation of~\cref{lem:lemma32} and suppose additionally that $2$ is invertible in $A$.
Then for $\ell \geq 7$ the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}
 Consider the following commutative diagram with exact rows, in which the lifting $t$ is obtained from~\cref{lem:lemma32}:
\begin{equation*} \xymatrix{
 C(\rD_\ell, B, I) \ar[r]^{j_B} \ar@{->>}[d]_{f} & \St(\rD_\ell, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\rD_\ell, B) \ar[r]^{\pi_B} \ar[d]_{h} & \St(\rD_\ell, k[X]) \ar[d]_{i} \\
 C(\rD_\ell, R, I) \ar[r]^{j_R}         & \St(\rD_\ell, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\rD_\ell, R) \ar[r]^-{\pi_R}        & \St(\rD_\ell, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{field-injectivity}, the element $a$
 also lies in $\Ker(\pi_B)$ and hence comes from some $b \in \St(\rD_\ell, B, I)$ via $\mu_B$.
Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\rD_\ell, R, I)$ such that $j_R(c) = g(b)$. 
By~\cref{lem:prop41}, $f$ is surjective, therefore $c = f(d)$ for some $d \in C(\rD_\ell, R, I)$ and we obtain the required assertion:
 \[ 1 = \mu_Bj_B(d) = tgj_B(d) = tj_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]

@@
